---
title: "Covid Death Rates, 1st and 2nd Differences"
author: "Brandon Tao"
date: "10/12/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 0. Setup

```{r message=FALSE}
library(tidyverse)
library(micromapST)
load('Data/Covid_Rates_Two_Week_Gaps.RData')
objects()
```


### 1. Arrow plots

Providing the labeling to describe a plot's content
in constrained spaced often poses a challenge. Shorter labels
can take up less space and simplify appearance. 
Here the units of measure label, 'Cases per 100,000' is
clear but long.  Below the label is shortened to 'Cases per 100K'.  

The grid line labels can also take up space. In this example 
The units are thousands so 4 digits long.  Rescaling the
units to percents is thinkable. Then there would be only one
digit each.  

Scaling Choices can be guided by the target audience. Those in the
health sciences are likely to think in terms of rates such
as rate per 100,000 or rate per million.  In criminology
rates per 1000, and 10,000 are likely more common.  Much of
public is more familiar with percents.   

The general strategy is have two or maybe three digits to the left of the
decimal place. This can motivate taking logs. Graphics and 
and modeling can also call for taking logs.     

Note that micromapST has some options in terms of gridline algorithms and
the alternating rows of gridline labels avoid overplotting.  

 

```{r fig.width=7.5, fig.height=10}


dat <- as.data.frame(st_deathRates3)
names(dat)

panelDesc1 <- data.frame(
  type = c('maptail','id','arrow','arrow','arrow'),
  lab1 = c('','','Sept. 11 and','Sept. 25 and','Sept. 11 and'),
  lab2 = c('','','Sept. 25','Oct. 9','Oct. 9 '),
  lab3 = c('','','Deaths per 100K','Deaths per 100K','Deaths per 100k'),
  col1 = c(NA,NA,'Sep_11_2020_R','Sep_25_2020_R','Sep_11_2020_R'), 
  col2 = c(NA,NA,'Sep_25_2020_R','Oct_09_2020_R','Oct_09_2020_R') 
)


micromapST(dat, panelDesc1,
  rowNamesCol = 'Postal',
  rowNames = 'ab',
  sortVar = 'Sep_11_2020_R', ascend = TRUE,
  title = c("Covid-19: Two-Week Death Rate Changes",
            "Fridays: Sept. 11 to Oct. 9, 2020 ")
)

```

### 2. Compute First and Second Differences  

For time series, second differences provide discrete analogues to the
second derivatives in continuous time series.  Trends usually do vary over
time.  With Covid-19 many variables drive the change in the trends.  We
don't know them all, and often don't have data on the variables that we
conjecture. Here we look at just one second difference.  

```{r}
names(st_deathRates3)
tib <- select(st_deathRates3,Postal,Sep_11_2020_R:Oct_09_2020_R)
names(tib)<- c('Postal','Fri1','Fri2','Fri3')
tibDif <- mutate(tib,
                dif1=Fri2-Fri1,
                dif2=Fri3-Fri2,
                dif3=dif2-dif1)
```

### 3. Show the differences using micromapST


```{r fig.width=7.5, fig.height=10}


dat <- as.data.frame(tibDif)
names(dat)

panelDesc2 <- data.frame(
  type = c('maptail','id','dot','dot','dot'),
  lab1 = c('','','Sept. 25 -','Oct. 9 -','Change in'),
  lab2 = c('','','Sept. 11','Sept. 25','Differences'),
  lab3 = c('','','Deaths per 100K','Deaths per 100K','Deaths per 100k'),
  col1 = c(NA,NA,'dif1','dif2','dif3')
)


micromapST(dat, panelDesc2,
  rowNamesCol = 'Postal', # Use this column
  rowNames = 'ab',  # Names are State Postal codes 
  sortVar = 'dif3', ascend = TRUE,
  title = c("First and Second Death Rate Differences",
            "Fridays: Sept. 11 to Oct. 9, 2020 ")
)

```