---
title: "Covid Case Rates, 1st and 2nd Differences"
author: "Brandon Tao"
date: "10/12/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 0. Setup

```{r message=FALSE}
library(tidyverse)
library(micromapST)
load('Data/Covid_Rates_Two_Week_Gaps.RData')
objects()
```


### 1. Arrow plots

Providing the labeling to describe a plot's content
in constrained spaced often poses a challenge. Shorter labels
can take up less space and simplify appearance. 
Here the units of measure label, 'Cases per 100,000' is
clear but long.  Below the label is shortened to 'Cases per 100K'.  

The grid line labels can also take up space. In this example 
The units are thousands so 4 digits long.  Rescaling the
units to percents is thinkable. Then there would be only one
digit each.  

Scaling Choices can be guided by the target audience. Those in the
health sciences are likely to think in terms of rates such
as rate per 100,000 or rate per million.  In criminology
rates per 1000, and 10,000 are likely more common.  Much of
public is more familiar with percents.   

The general strategy is have two or maybe three digits to the left of the
decimal place. This can motivate taking logs. Graphics and 
and modeling can also call for taking logs.     

Note that micromapST has some options in terms of gridline algorithms and
the alternating rows of gridline labels avoid overplotting.  

 

```{r fig.width=7.5, fig.height=10}


dat <- as.data.frame(st_caseRates3)
names(dat)

panelDesc1 <- data.frame(
  type = c('maptail','id','arrow','arrow','arrow'),
  lab1 = c('','','Sept. 11 and','Sept. 25 and','Sept. 11 and'),
  lab2 = c('','','Sept. 25','Oct. 9','Oct. 9 '),
  lab3 = c('','','Cases per 100K','Cases per 100K','Cases per 100k'),
  col1 = c(NA,NA,'Sep_11_2020_R','Sep_25_2020_R','Sep_11_2020_R'), 
  col2 = c(NA,NA,'Sep_25_2020_R','Oct_09_2020_R','Oct_09_2020_R') 
)


micromapST(dat, panelDesc1,
  rowNamesCol = 'Postal',
  rowNames = 'ab',
  sortVar = 'Sep_11_2020_R', ascend = TRUE,
  title = c("Covid-19: Two-Week Case Rate Changes",
            "Fridays: Sept. 11 to Oct. 9, 2020 ")
)

```

Describing the plots will be address in class. The focus here is 
on plot production.
  
I often mistakes in the panel description dataframe.  The micromapST function
provides feedback that can be helpful in identifying problems. When using an R
script the feedback is likely to appear on Console.  When using RMD files 
some feedback appears little panel below the script chuck.  There is very little yellow
warning triangle in the panel. Clicking on the panel reveals messages.   
  
For this example, the message I saw indicated that variables specified were not found.
That was a good clue. I had typed lower case r's and not upper case R's in the
variables names.  
  
On occasion I misspell the sorting variable name. Then, with no other problems, 
micromapST will produce the plot without sorting, and provide a warning message. 
The lack of sorting is obvious in the plot but sometimes people in a hurry
and assume the plot produced is what they intended.

### 2. Compute First and Second Differences  

For time series, second differences provide discrete analogues to the
second derivatives in continuous time series.  Trends usually do vary over
time.  With Covid-19 many variables drive the change in the trends.  We
don't know them all, and often don't have data on the variables that we
conjecture. Here we look at just one second difference.  

```{r}
names(st_caseRates3)
tib <- select(st_caseRates3,Postal,Sep_11_2020_R:Oct_09_2020_R)
names(tib)<- c('Postal','Fri1','Fri2','Fri3')
tibDif <- mutate(tib,
                dif1=Fri2-Fri1,
                dif2=Fri3-Fri2,
                dif3=dif2-dif1)
```

### 3. Show the differences using micromapST


```{r fig.width=7.5, fig.height=10}


dat <- as.data.frame(tibDif)
names(dat)

panelDesc2 <- data.frame(
  type = c('maptail','id','dot','dot','dot'),
  lab1 = c('','','Sept. 25 -','Oct. 9 -','Change in'),
  lab2 = c('','','Sept. 11','Sept. 25','Differences'),
  lab3 = c('','','Cases per 100K','Cases per 100K','Cases per 100k'),
  col1 = c(NA,NA,'dif1','dif2','dif3')
)


micromapST(dat, panelDesc2,
  rowNamesCol = 'Postal', # Use this column
  rowNames = 'ab',  # Names are State Postal codes 
  sortVar = 'dif3', ascend = TRUE,
  title = c("First and Second Case Rate Differences",
            "Fridays: Sept. 11 to Oct. 9, 2020 ")
)

```